---
title: "Data Analysis Using R"
output:
  html_document: default
  pdf_document: default
---

## Learning Outcomes
* Ability to organize, visualize and analyze data by statistical software R.
* Skills to select and use an appropriate statistical method to solve the real-world problem.
* Report writing and oral presentation skills
* Develop ability on writing R program to solve various real-life problem.

## Topics：
* Chapter One: Introduction to `R`
* Chapter Two: From data to graphics
* Probability Distributions
* Hypothesis Testing and Confidence Interval Estimation
* Analysis of variance
* Multiple Regression
* Tabular Data

## Leason Resource

### Introduction to `R`

#### What is `R`
`R` is a programming language with is **widely** using in statistics and data science.
- software;
- documentation (R-Markdown and Jupyter-notebook);
- RNews.

#### How to install `R`
<a href="https://cran.r-project.org">**Click here**</a> and you can find the answer. \\ **Remember!**, follow the direction and you may get the `R` installed properly.

#### Some Useful Commands

* How to Quit the `R` program
`> q()`
```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
q() # attention! DO NOT RUN THIS CODE IN THIS MOMENT
```
If you are coding in Windows, `File_Exit` can also works.

* Get or Set Working Directory(Files Path)
`getwd` returns an absolute filepath representing the current working directory of the `R` process; `setwd(dir)` is used to set the working directory to dir.
```{r}
getwd()
setwd(dir)
```
* List Objects
`ls` and `objects` return a vector of character strings giving the names of the objects in the specified environment. When invoked with no argument at the top level prompt, `ls` shows what data sets and functions a user has defined. When invoked with no argument inside a function, `ls` returns the names of the function's local variables: this is useful in conjunction with browser.
```{r}
ls()
objects()
```
