---
title: "Assignment 3"
author: "Terry Liu"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question One

The datat file **Real-Estate.txt** contains information on the homes sold in the Denver area during the yea 2003. The variables in this data file are as follows: 

| Name | Representation |
| ---- | ------------ |
|Price|Selling price in $1000|
|Bedrooms|Number of bedrooms|
|Size|Size of the home in square feet|
|Pool|Swimming Pool(1=Yes,0=No)
|Distance|Distance from the home to the center of the city|
|Township|Township NO.|
|Garage|Garage attached(1=Yes,0=No)|
|Baths|Number of bathrooms|

### Part A
Give a plot of selling price agaignst distance from the home to the center of the city. Does there seem to be a relationship between the two variables? If so, is the relationship direct or inverse ? 

### Answer: 
```{r cars}
db <- read.table("Real-Estate.txt",header=TRUE)

x <- db$Distance
y <- db$Size

plot(x,y)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
