---
title: "Regression_Analysis"
author: "Terence Lau"
date: "11/25/2018"
output: word_document
---

## Get Ready

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Beijing_Data <- read.csv("PRSA_data_2010.1.1-2014.12.31.csv", header = T)
# No: row number 
# year: year of data in this row 
# month: month of data in this row 
# day: day of data in this row 
# hour: hour of data in this row 
# pm2.5: PM2.5 concentration (ug/m^3) 
# DEWP: Dew Point (℃) 
# TEMP: Temperature (℃) 
# PRES: Pressure (hPa) 
# cbwd: Combined wind direction 
# Iws: Cumulated wind speed (m/s) 
# Is: Cumulated hours of snow 
# Ir: Cumulated hours of rain 
```

## Data Summary 

```{r}
head(Beijing_Data)
tail(Beijing_Data)
```

There are $12$ variables in this data set: 
| Variable Name | Variable Explanation  |
| :-----------: | :-------------------: |
| No | row number |
| Year | year of data in this row |
| Month | month of data in this row |
| Day | day of data in this row |
| Hour | hour of data in this row |
| PM 2.5 | PM 2.5 concentration (ug/m^3) |
| DEWP | Dew Point |
| TEMP | Temperature |
| PRES | Pressure |
| cbwd | Combined wind direction | 
| lws | Cumulated wind speed | 
| ls | Cumulated hours of snow |
| lr | Cumulated hours of rain |

We want to find out a model to analyze this data set and predict the fulture data. 

## Data Preprocessing 

### Handling Missing Values 

First, we have to check if there are missing values,

```{r}
length(Beijing_Data[is.na(Beijing_Data) == T])
```

There are $2067$ missing values out of all $434824$ obervations of PM2.5 values

```{r}
beijing$pm2.5 <- impute(beijing$pm2.5, mean) # missing value fill up
```




