---
title: "Assignment 3"
author: "Terry Liu"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question One

The datat file **Real-Estate.txt** contains information on the homes sold in the Denver area during the yea 2003. The variables in this data file are as follows: 

| Name | Representation |
| ---- | ------------ |
|Price|Selling price in $1000|
|Bedrooms|Number of bedrooms|
|Size|Size of the home in square feet|
|Pool|Swimming Pool(1=Yes,0=No)
|Distance|Distance from the home to the center of the city|
|Township|Township NO.|
|Garage|Garage attached(1=Yes,0=No)|
|Baths|Number of bathrooms|

### Part A
Give a plot of selling price agaignst distance from the home to the center of the city. Does there seem to be a relationship between the two variables? If so, is the relationship direct or inverse ? 

### Answer: 
```{r}
db <- read.table("Real-Estate.txt",header=TRUE)
names(db)

x <- db$Distance
y <- db$Price

plot(x,y)
abline(lm(y~x))
```
We can have the conclusion that there is negative relationship between distance and the selling prze.

### Part B

Plot a histogram for selling price. Comment on the shape of the distribution. 
```{r}
hist.default(db$Price,xlim=c(50,400),main='Price')

skewness <- function (x, na.rm = FALSE) 

{

  if (na.rm) 

    x <- x[!is.na(x)]

  sum((x - mean(x))^3)/(length(x) * sd(x)^3)

}


skewness(db$Price)

```
We can get the skewness of these data (Selling Price) is equals to $0.4605562$, so, it is Positive Skew. 

### Part C 

Create a side-by-side box plots to compare the distribution of selling price for homes with an attached garage and homes without a garage. Comment on what you observe.

```{r}
x_1 <- data1$Price[data1$Garage==1]
x_2 <- data1$Price[data1$Garage==0]
Garage <- c("with Garage","without Garage")
boxplot(x_1,x_2,names = Garage, horizontal = TRUE, main = "The Price comparison",xlab="Price")
```
Observation: The averge price of those have garage are higher than without garage.

### Part D 

There are five Townships in htis data. Draw a bar plot to compare the number of homes sold in these Townships. 

```{r}
township_1 <- length(data1$Township[data1$Township == 1])
township_2 <- length(data1$Township[data1$Township == 2])
township_3 <- length(data1$Township[data1$Township == 3])
township_4 <- length(data1$Township[data1$Township == 4])
township_5 <- length(data1$Township[data1$Township == 5])

name_township <- c("township1", "township2", "township3", "township4", "township5")
township <- c(township_1, township_2, township_3 ,township_4, township_5)
tmp <- as.vector(name_township)  

bar <- barplot(township, axes=T, names = name_township, las=2)


# axis(side = 1, lwd = 1, tck = -.015, labels = NA)
# # 添加x轴的坐标轴(包括坐标轴，刻度线和刻度线标签)，但是却未显示刻度线标签
# axis(side = 2, lwd = 1, tck = -.015, labels = NA) # 添加y轴的坐标轴(包括坐标轴，刻度线和刻度线标签)，但是却未显示刻度线标签
# axis(side = 1, lwd = 0, line = -.4) # 第二次添加x轴的坐标轴(包括坐标轴，刻度线和刻度线标签)，但是只显示了刻度线标签(通过lwd = 0使坐标轴，刻度线未显示出来)
# axis(side = 2, lwd = 0, line = -.4, las = 1) # 第二次添加y轴的坐标轴(包括坐标轴，刻度线和刻度线标签)，但是只显示了刻度线标签
        
        
```

### Part E 

Draw a pie chart of the varialbe "Baths" and label the percentage of each number of Bathrooms 

```{r}
x <- c(unique(data1$Baths))

length_x <- length(x)



lbls <- x
pct <- round(x/sum(x)*100)
lbls <- paste(lbls, pct) # add percents to labels 
lbls <- paste("Baths=",lbls,"%",sep="") # ad % to labels 
pie(x,labels = lbls, col=rainbow(length(lbls)),
  	main="Pie Chart of Baths")

```

## Question Two

Recently, it has been proposed that there is a non-linear relation between the number of perople $N$ in a city, and its economic output per person $Y$ (per captia \gross metropolitan product): 
$$Y = y_0N^a + \text{ noise }$$


Using data from the 366 metropolitan area of the US (as of 2006) to estimate $a$, by minimizing the mean squared error,
$$\frac{1}{366} \sum_{i=1}^366 (Y_i - y_0N_i^a)^2$$
over all a. We will ess later in the class how we could estimate both parameters at onece, but for now we'll treat $y_0$ as know, and in particular equal = $6611$.






































