---
title: "Data Analysis Using R"
output:
  html_document: default
  pdf_document: default
---

## Learning Outcomes
* Ability to organize, visualize and analyze data by statistical software R.
* Skills to select and use an appropriate statistical method to solve the real-world problem.
* Report writing and oral presentation skills
* Develop ability on writing R program to solve various real-life problem.

## Topics：
* Chapter One: Introduction to `R`
* Chapter Two: From data to graphics
* Probability Distributions
* Hypothesis Testing and Confidence Interval Estimation
* Analysis of variance
* Multiple Regression
* Tabular Data

## Leason Resource

### Introduction to `R`

#### What is `R`
`R` is a programming language with is **widely** using in statistics and data science.
- software;
- documentation (R-Markdown and Jupyter-notebook);
- RNews.

#### How to install `R`
<a href="https://cran.r-project.org">**Click here**</a> and you can find the answer. \\ **Remember!**, follow the direction and you may get the `R` installed properly.

#### Some Useful Commands

* How to Quit the `R` program
`> q()`
```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
q()
# attention! DO NOT RUN THIS CODE IN THIS MOMENT
# quit(save = "default", status = 0, runLast = TRUE)
# q(save = "default", status = 0, runLast = TRUE)
```
If you are coding in Windows, `File_Exit` can also works.

* Get or Set Working Directory(Files Path)
`getwd` returns an absolute filepath representing the current working directory of the `R` process; `setwd(dir)` is used to set the working directory to dir.
```{r}
getwd()
setwd(dir)
```
* List Objects
`ls` and `objects` return a vector of character strings giving the names of the objects in the specified environment. When invoked with no argument at the top level prompt, `ls` shows what data sets and functions a user has defined. When invoked with no argument inside a function, `ls` returns the names of the function's local variables: this is useful in conjunction with browser.
```{r}
ls()
# ls(name, pos = -1L, envir = as.environment(pos),all.names = FALSE, pattern, sorted = TRUE)
objects()
# objects(name, pos= -1L, envir = as.environment(pos),all.names = FALSE, pattern, sorted = TRUE)
```
#### Getting HELP
* Details about a specific command whose name you know (input arguments, options)
`help()` is the primary interface to the help systems
```{r}
help()
#help(topic, package = NULL, lib.loc = NULL,verbose = getOption("verbose"),try.all.packages = getOption("help.try.all.packages"),help_type = getOption("help_type"))
```
Also, we can use `?` + `[command]` to get the help files, for instance, `?ls()`

* Find commands containing a regular expression:
`apropos()` returns a character vector giving the names of objects in the search list matching (as a regular expression) what.
```{r}
apropos("help")
# The return:
# [1] "help"          "help.request"  "help.search"   "help.start"    "main.help.url"
```
* Web Searching
We can also call the search engine lets us to search for topic related to regular expressions:
`>help.search("help")`This command allows for searching the help system for documentation matching a given character string in the (file) name, alias, title, concept or keyword entries (or any combination thereof), using either fuzzy matching or regular expression matching. Names and titles of the matched help entries are displayed nicely formatted.
```{r}
help.search("covariance")
```
* Example of Usage
`demo` is a user-friendly interface to running some demonstration R scripts. demo() gives the list of available topics.
```{r}
demo(graphics)
#demo(topic, package = NULL, lib.loc = NULL,character.only = FALSE, verbose = getOption("verbose"),echo = TRUE, ask = getOption("demo.ask"),encoding = getOption("encoding"))
```
* Check the function
To see code for a function, just type the name with no parentheses or arguments \\
```{r}
plot
function (x, y, ...)
UseMethod("plot")
<bytecode: 0x7fb062f6f9d8>
<environment: namespace:graphics>
```
#### `R` As A Calculator
We can use `R` to do some calculation just using the operators: `+, -, *, /` and etc..
```{r}
2 + 2
# [1] 4
2 * 3 * 4 * 5
# [1] 5
log2(32)
# [1] 5
print(sqrt(2))
# [1] 1.414214
sin(c(30,60,90) * pi / 180)
# [1] 0.5000000 0.8660254 1.0000000
```
#### Important concept

* Command:function
```{r}
plot(mydata1),q()
#mydata1 is not exsist
```
* Objects:
vector, matrix, factor, list and data frame

* Modes
logical, numeric, and character

#### Modes
* Variables:
```{r}
a <- -49 # let 'a' equals negative fourty-nine
sqrt(a) # solution is: [1] 7
```
`a` is numeric.

``` {r}
b <- "The dog ate my homework"
sub("dog", "cat", b)
# substitude "dog" with "cat"
#solution is: [1] "The cat ate my homework"
```
`b` is Character string

```{r}
d <- (1 + 1 == 3)
d # solution is: [1] FALSE
```
`d` is Logical(Boolean Variable)

#### Objects
* `ls()` and `objects` return a vector of character strings giving the names of the objects in the specified environment.
* `rm()` and `remove` can be used to remove objects.
```{r}
vector_1 <- c(10,20,30)
rm(vector_1)
vector_1
# solution is: Error: object 'vector_1' not found
```
* `typeof` determines the (R internal) type or storage mode of any object
```{r}
> tmp <- 1:4
> typeof(tmp)
# solution is: [1] "integer"
```
* `class()` a simple generic function mechanism which can be used for an object-oriented style of programming.
```{r}
x <- 10
class(x)
# solution is [1] "numeric"
```
#### Vectors

The concept of Vector is vitally important in every area of maths and the kye feature which makes `R` very useful for statistics is that it is vectorized. This means thay many operations can be performed point-wise on vector. `c()` is the function to create vector:
We can check the R documentation: `c()` is a generic function which combines its arguments.
```{r}
c(...)## S3 Generic function

c(..., recursive = FALSE, use.names = TRUE)## Default S3 method
```
```{r}
x <- c(1, -1, 3.5, 2)
print(x) # solution is: [1]  1.0 -1.0  3.5  2.0

x + 2 # add 2 to every number in this vectors
print(x + 2) # solution is: [1] 3.0 1.0 5.5, 4.0

x ^ 2 # square each entry
print(x ^ 2) # solution is: [1]  1.00  1.00 12.25  4.00
```
Also, we can generate integer sequences by different method(colon operator):
```{r}
1:10 #solution is: [1]  1  2  3  4  5  6  7  8  9 10
-3:4 #solution is: [1] -3 -2 -1  0  1  2  3  4
9:4 #solution is: [1] 9 8 7 6 5

seq(from=2, to=6, by=0.4) #solution is: [1] 2.0 2.4 2.8 3.2 3.6 4.0 4.4 4.8 5.2 5.6 6.0

#`seq` is a standard generic with a default method.
## Default S3 method:
#seq(from = 1, to = 1, by = ((to - from)/(length.out - 1)),length.out = NULL, along.with = NULL, ...)
```
Sometimes it's necessary to have repeated values, for which we use `rep()`
```{r}
rep(5,3) # solution is: [1] 5 5 5
rep(2:5, each=3) # solution is: [1] 2 2 2 3 3 3 4 4 4 5 5 5
rep(-1:3, length.out=10) # solution is: [1] -1 0 1 2 3 -1 0 1 2 3
```
We can also use R's vectorization to create more interesting sequences:
```{r}
2^(0:10) # solution is: [1] 1 2 4 8 16 32 64 128 256 5121024
1:3 + rep(seq(from=0,by=10,to=30), each=3) # solution is: [1] 1 2 3 11 12 13 21 22 23 31 32 33
```
It's frequently necessary to extract some of the elements of a larger **vector**.
In `R` you can use square brackets to select an individual element or group of elements:
```{r}
x <- c(5,9,2,14,-4)
x[3]
#solution is: [1] 2

x[c(2,3,5)] # note indexing starts from 1
# solution is: [1] 9 2 -4

x[1:3] # print out from 1 to 3
# solution is: [1] 5 9 2

x[3:length(x)]
print(length(x)) #length(x) can gets or sets the length of vectors
print((x[3:length(x)]) == x[3:5])
```
There are two other methods for getting subvectors. The first is using a logical vector ( i.e. containing `TRUE` and `FALSE` of the same length:)
```{r}
x > 4
# solution is: [1] TRUE TRUE FALSE TRUE FALSE
x[x > 4]
# solution is: [1] 5 9 14
```
or we can use negative indices to specify wich elements should not be selected:
```{r}
x[1]
# solution is [1] 9 2 14 -4
x[-c(1,4)]
# solution is 9 2 -4
```
(Note that this is rather different tot what other languages such as `C` or `Python` would interpret negative indices to mean.)

* Logical Operators
As we see above, the comparison operator `>` returns a logical vector indicating whether or not the left hand side is greater that the right hand side. Here we demonstrate the other comparison operators:
 ```{r}
 x <= 2 # less than or equal to
 # solution is: [1] FALSE FALSE TRUE FALSE TRUE
 x == 2 # equal to
 # solution is: [1] FALSE FALSE TRUE FALSE FALSE
 x != 2 # not equal to
 # solution is: [1]  TRUE  TRUE FALSE  TRUE  TRUE
```
Notice that the double equals sign `==`, to distinguish between assignment and comparison.
We may also wish to combine logical vectors. If we want the elements of x within a gange, we can use the following:
```{r}
(x > 0) & (x<10) # '&' means 'and'
# solution is: [1] TRUE TRUE TRUE FALSE FALSE

# The & operator does a pointwise 'and' comparison between the two sides. Similarly, the vertical bar | does pointwise 'or', and the unary `!` operator performs negation.

(x == 5) | (x > 10)
# solution is: [1]  TRUE FALSE FALSE  TRUE FALSE
!(x > 5)
# solution is: [1]  TRUE FALSE  TRUE FALSE  TRUE
```
* Character Vectors
As you might have noticed in the above section, vectors only have contains numbers, but vectors can also contain numbers. We can equally create a **Character Vectors**, in which each entry is a string of text. Strings in `R` are contained within double quote":
```{r}
x <- c("Hello", "how do you do", "lovely to meet you", 42)
print(x)
# the solution is :
# [1] "Hello"              "how do you do"      "lovely to meet you"
# [4] "42"
```
Notice that you cannot mix numbers with strings: if you try to do so the number will be converted into a string. Otherwise charater vectors are much like their numerical counterparts.
```{r}
x[2:3]
# solution is: [1] "how do you do" "lovely to meet you"
x[-4]
$ solutions is
# solution is: [1] "Hello" "how do you do" "lovely to meet you"
c(x[1:2], "goodbye")
# solution is: [1] "Hello"  "how do you do" "goodbye"
```
#### Matrices
* Matrix: rectangular table of data of the sam type (Two dimension)
* A mtrix:
```{r}
matrix(1:10, ncol = 5)`
#      [,1] [,2] [,3] [,4] [,5]
# [1,]    1    3    5    7    9
# [2,]    2    4    6    8   10
# matrix function Description
# ‘matrix’ creates a matrix from the given set of values.
# ‘as.matrix’ attempts to turn its argument into a matrix.
# ‘is.matrix’ tests if its argument is a (strict) matrix.

# Usage:
# matrix(data = NA, nrow = 1, ncol = 1, byrow = FALSE, dimnames = NULL)
```
**Create Matrix**
We can use `matrix` function to create a matrix, the usage can check the help.
```{r}
matrix(1:9, nrow=3, ncol=3)
#       [,1] [,2] [,3]
# [1,]    1    4    7
# [2,]    2    5    8
# [3,]    3    6    9

matrix(1:9, nrow=3, ncol=3, byrow=T) #byrow = T means tranfer row and col
#      [,1] [,2] [,3]
# [1,]    1    2    3
# [2,]    4    5    6
# [3,]    7    8    9
